Quantifying passive kelp rafting phenomena and connectivity of nearshore communities around the coast of South Africa through development of a 3D spectral numerical coastal-flow model

### 2020-07-23 thesis_chapter_1.tex
added heading

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Quantifying passive kelp rafting phenomena and connectivity of nearshore communities model around the coast of South Africa:Development of a 3D spectral numerical coastal-flow},
  pdfauthor={RM Coppin; C Rautenbach; AJ Smit},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% https://github.com/rstudio/rmarkdown/issues/337
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

@@

\title{Quantifying passive kelp rafting phenomena and connectivity of nearshore
communities model around the coast of South Africa:Development of a 3D
spectral numerical coastal-flow}
\author{RM Coppin\footnote{University of the Western Cape} \and C Rautenbach\footnote{South African Weather Service} \and AJ Smit\footnote{University of the Western Cape}}
\date{}

@@

\begin{figure}

{\centering \includegraphics{thesis_chapter_1_files/figure-latex/surface elevation figures-1}

}

\caption{\label{fig:surface elevation figures}The definition of a wave in a time record of the surface elevation with downward zero-crossings (upper panel) and upward zero-crossings (lower panel).}\label{fig:surface elevation figures}
\end{figure}

@@

\begin{figure}

{\centering \includegraphics{thesis_chapter_1_files/figure-latex/Hs_Tp_surface_elevation figure-1}

}

\caption{The definition of wave height and wave period in a time record of the surface elevation (the wave is defined with downwrd zero-crossings.}\label{fig:Hs_Tp_surface_elevation figure}
\end{figure}

@@

@@

The trajectory of passively drifting objects on the sea surface is
influenced by multiple factors, such as water currents, atmospheric
wind, wave motion, wave induced currents, gravitational force and
buoyancy force. To complicate matters, the previously mentioned factors
do not act independently of one another but instead influence one
another. Furthermore, the gravitational and buoyancy forces on the
object are determined by the objects shape. Therefore, all these factors
need to be taken into account when modelling trajectory of drifting
objects. Given the local wind, surface current, and the shape and
buoyancy of the object is known, it is possible to estimate trajectory
by the equation \[V_{drift}=V_{curr} + V_{rel}\]

where \(V_{curr}\) represents the current velocity relative to the
earth, and \(V_{rel}\) represents the object drift velocity relative to
the ambient water. Ocean currents are determined by two components: the
surface current (including the effects of Ekman drift, baroclinic
motion, tidal and inertial currents) and Stokes drift induced by waves.
The assumption made is \(V_{curr}\) influences all floating objects in
the same manner, regardless of shape or size. Therefore, \(V_{curr}\) is
equated with the surface current obtained from a numerical ocean model
which has been parameterised by wind velocity. The effects of
\(V_{rel}\) on a floating object is driven by wind and wave forces which
is dependent on the shape and size of the floating object. Previous
research has shown that the effects of waves on drifting objects are
negligible when the length of the object is less than the wave length;
and increase significantly when the lengths are approximately the same
(Grue and Biberg 1993; Hodgins and Hodgins 1998). Furthermore, current
velocity \(V_{curr}\) affects on drifting objects is considered to be
negligible in this approach and ideally should be taken into account.
Studies investigating kelp rafts have identified surface current
velocity (as a function of wind velocity) as an important influence in
determine trajectory (Harrold and Lisin 1989; Macaya et al. 2005; Hobday
2000a, 2000b). Therefore, the objects motion/trajectory relative to the
wind must be taken into account.

\hypertarget{windage-factors}{%
\subsection{Windage factors}\label{windage-factors}}

An objects motion relative to the wind is sometimes referred to as
``windage factors'' or "leeway drift, and is difficult to accurately and
empirically describe. The difficulty is due to accurate measurements
needed for the current velocity, wind velocity, wave height and wave
direction. Furthermore, these factors do not act independently and
instead influence one another, which compounds the complexity of
calculating wind effects on drifting objects. To overcome this
challenge, measurements of current and wind velocity, and wave height
and direction are measured along the objects trajectory for a wide range
of conditions.

@@

Important to note is that the effects of wind and waves on the
trajectory of a drifting object may be influenced by the object's
orientation. This has been noted in modelling drift trajectories of
ships (Allen and Plourde 1999; Hackett, Breivik, and Wettre 2006), and
should be incorporated into the leeway drift component. The direct
(\(V_l\)) and indirect (\(V_s\) and \(V_w\)) effects do not act
independently and instead influence one another. with regards to this,
Griffin, Oke, and Jones (2017) recommend incorporating an effective
windage factor which would represent the combined effects of Stokes
Drift, leeway drift and wave forces. Although a common approach, this
may not be necessary in all circumstances particularly if individual
effects of factors. As mentioned previously, the magnitude of effects of
waves and wind on a drifting object is dependent on its shape, size and
buoyancy.

@@

@@

\begin{figure}

{\centering \includegraphics{thesis_chapter_1_files/figure-latex/BF_1-1}

}

\caption{\label{fig:BF_1}Conceptual model of bouyancy characteristics.}\label{fig:BF_1}
\end{figure}

@@

\hypertarget{buoyancy-factors}{%
\subsection{Buoyancy factors}\label{buoyancy-factors}}

The buoyancy of objects at sea has been shown to be influeced by
epibiont load (Hobday 2000b; Rothä usler et al. 2011; Graiff et al.
2016; Macaya et al. 2016), which reduces drift times by reducing
buoyancy with increasing biomass, which ultimatley leads to the object
sinking. The bouyancy of the drifting object is determined by the growth
rate of epiphytic species while drifting, which in turn will be depndent
on environmental factors such as temperature and light which will vary
along the objects trajectory Therefore, buoyancy should be parameterised
by temporal epibiont biomass load.

\hypertarget{particle-dispersion-modelling}{%
\section{Particle dispersion
modelling}\label{particle-dispersion-modelling}}

\hypertarget{particle-dispersion-models}{%
\subsection{Particle dispersion
models}\label{particle-dispersion-models}}

The Pol3DD model is able to track virtual particles to simulate water
borne dispersion of material. Examples include neutrally buoyant
anthropogenic material, larvae, oil spills, outfall discharges and
sediment transport (Lebreton, Greer, and Borrero 2012). The model uses a
second-order accurate advection scheme (Black and Gay 1990) described as
follows:
\[\delta\chi=\frac{({u'+ \frac{(u_y\upsilon'-\upsilon_yu')\delta t}{2} )\delta t}}{(1-\frac{u_x\delta t}{2})(1-\frac{\upsilon_y\delta t}{2})-\frac{u_y\upsilon_\chi\delta t^2}{4}}\]

\[\delta y=\frac{({\upsilon' + \frac{(\upsilon_\chi u'-u_\chi\upsilon) \delta t}{2})\delta t}}{(1-\frac{u_y\delta t}{2})(1-\frac{\upsilon_\chi\delta t)}{2}) - \frac{u_\chi\upsilon_y\delta t^2}{4}}\]
where

\[u'= u + \frac{\delta tu_t}{2}\]
\[\upsilon'=\upsilon + \frac{\delta t\upsilon_t}{2}\] and \(u\),
\(\upsilon\) are orthogonal velocity components, \(t\) is time,
\(\delta t\) is the model time step, \(u_\chi\), \(u_y\), \(\upsilon_y\)
are the \(u\) and \(\upsilon\) spatial velocity gradients, and \(u_t\),
\(\upsilon_t\) are temporal gradients. Horizontal diffusion was modeled
as a random walk with separate longitudinal and lateral coefficients set
to simulate random turbulence. The distance increments moved by the
particle at each time step is calculated as:
\[\Delta X = R_N \sqrt{6E_1\delta t}\]
\[\Delta y = R_N\sqrt{6E_2\delta t}\] where \(R_N\) is a random number
in the uniform range (-1,1) and \(E_1,E_2\) are the longitudinal and
lateral eddy diffusivities, respectively. To estimate the amount of
input into the model the authors used a scaled approach to define the
release of particles, and where based on previous research performed by
Halpern et al. (2008).

\hypertarget{application-of-particle-dispersion-models}{%
\section{Application of particle dispersion
models}\label{application-of-particle-dispersion-models}}

A study by Lebreton, Greer, and Borrero (2012) proposed a methodology to
track floating debris from source to sink based on descriptions of
global waste production and ocean surface currents, using a combination
of hydrodynamic and particle dispersion modelling. The particle tracking
model applied uses a two stage process, (1) a hydrodynamic model solves
the equations of motion to characterise the movement of water within the
domain of the model, (2) virtual particles are released into the flow
field established in the first step, and particles are allowed to move
through hydrodynamic forcing. The authors extracted sea surface currents
from the HYCOM/NCODA ocean circulation modelling system, which is forced
by the US Navy's Operational Global Atmospheric Prediction System
(NOGAPS). The model also took into account wind stress, wind speed, heat
flux, and precipitation and was looped five times to accurately
represent ocean circulation patterns for a 30 year period. The authors
used the velocity data aquired from HYCOM and coupled it to the
Lagrangian particle tracking model Pol3DD which was used to simulate
dispersion. No additional wind stress terms were applied to the motion
of particles, as this factor had already been incorporated in the HYCOM
hydrodynamic data. Potential extra stress for emerged parts of debris
was considered negligible as all floating debris are assumed to be fully
submerged. The Pol3DD model was also modified so provide additional
information on individual particles such as origin, age, and trajectory
information of individual particles. Since the model considered a
spherical earth, particles were allowed to re-enter the simulation if
they passed around the `globe'. Important to note is only the surface
layer was modelled in this study, as much of the evidence identifies
surface currents as the main dispersal vector of anthropogenic debris.

Griffin, Oke, and Jones (2017) added to the search for flight MH370 by
providing estimates of the trajectory of the debris. This was achieved
through a combination of an advanced ocean model calibrated with oceanic
drifters, and drifters in the form of aircraft parts. Unlike previous
work regarding the recovery of MH370 wreckage, this study incorporated
windage factors into the trajectory model as well as providing better
estimates of Stokes Drift.

@@

Dispersal is recognised as important driver of biodiversity, composition
and structure of ecological systems in the marine environment (Bernardes
Batista et al. 2018; {\textbf{???}}; Highsmith 1985; MacArthur and
Wilson 2001; Jackson and Sax 2010). These processes are largely
dependent on currents for dispersal of migrant populations which
ultimately promote connectivity of marine ecosystems (Bernardes Batista
et al. 2018; Mackas, Denman, and Abbott 1985, 1985; Zakas et al. 2009).
This is particularly true for organisms which lack pelagic larvae, and
are reliant on other modes of transport to reach new habitats through
passive modes of dispersal from anthropogenic and natural sources.
Anthropogenic sources include marine litter and structural debris while
examples of organic sources include pumice and macroalgae in the form of
kelp-rafts. Kelp-rafts have been identified as an important mode of
passive dispersal of marine organisms, such as invertebrates
{[}({\textbf{???}}); whichmann2012{]}, epiphytes {[}@{]}, grazers
(Nikula et al. 2010), invasive species (Lewis, Riddle, and Smith 2005)
and macroalgae themselves (Macaya et al. 2005) which either make
landfall or are transported offshore where they eventually sink. A
kelp-raft consists of an entanglement of macroalgae (one or multiple
species) which have been dislodged from the benthic environment through
hydrodynamic forces (mostly storms) and are positively buoyant by means
of gas-filled pneumatocysts and/or reproductive organs. Kelp-rafts are
capable of travelling vast distances {[}@{]} and are considered
important dispersal vectors in temperate latitudes, such as the Southern
California Bight (Hobday 2000a) and the Northern Baltic Sea {[}@{]}.

Kelp-raft abundance has been shown to vary temporally and spatially in
the ocean and around coastlines, which is dependent on seasonal growth
patterns. For example, studies investigating the dispersal patterns of
\emph{Sargassum} in the West Pacific show increased abundance during
growth seasons (spring and summer) as more biomass is available to
fragment (Deysher and Norton 1981; Kingsford 1992). Temporal variability
may also be related to seasonal storm frequency where more storms in a
particular season relate to an increase in kelp-raft abundance. For
example a study by Kingsford (1995) investigated the contribution of
\emph{Macrocystis pyrifera} rafts to habitat complexity in pelagic
environments. The authors found that the abundance of \emph{M. pyrifera}
rafts increased during seasons where storms occurred more frequently.

The trajectory of kelp rafts are determined by a combination of
oceanographic conditions and biological processes. Since kelp rafts
float, evidence points largely to wind driven surface currents as the
main driver, although the relative importance of wind versus surface
currents is largely unknown. For example a study by Harrold and Lisin
(1989) used radio-trackers on both natural and artificial
\emph{Macrocystis pyrifera} kelp rafts to investigate seasonal
trajectory patterns. The results showed that tracked kelp rafts were
mostly deposited nearby the source population and that seasonal wind
direction was the primary driver of trajectory for kelp rafts submerged
roughly \textasciitilde0.5 - 1m below the surface Harrold and Lisin
(1989). However, the authors did note the influence of surface currents
for some of their experiments when winds relaxed; which suggests that
surface currents may play a larger role in determining trajectory for
kelp rafts less exposed to wind or when prevailing winds are low.

@@

\leavevmode\hypertarget{ref-black1990}{}%
Black, Kerry P, and Stephen L Gay. 1990. ``A Numerical Scheme for
Determining Trajectories in Particle Models,'' 151--56.

@@

@@

\leavevmode\hypertarget{ref-halpern2008}{}%
Halpern, Benjamin S, Shaun Walbridge, Kimberly A Selkoe, Carrie V
Kappel, Fiorenza Micheli, Caterina D'agrosa, John F Bruno, et al. 2008.
``A Global Map of Human Impact on Marine Ecosystems.'' \emph{Science}
319 (5865): 948--52.

@@

@@

\leavevmode\hypertarget{ref-lebreton2012}{}%
Lebreton, LC-M, SD Greer, and Jose Carlos Borrero. 2012. ``Numerical
Modelling of Floating Debris in the World's Oceans.'' \emph{Marine
Pollution Bulletin} 64 (3): 653--61.

@@

\leavevmode\hypertarget{ref-rotha2011}{}%
Rothä usler, Eva, Iván Gómez, Iván A Hinojosa, Ulf Karsten, Leonardo
Miranda, Fadia Tala, and Martin Thiel. 2011. ``Kelp Rafts in the
Humboldt Current: Interplay of Abiotic and Biotic Factors Limit Their
Floating Persistence and Dispersal Potential.'' \emph{Limnology and
Oceanography} 56 (5): 1751--63.

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Quantifying passive kelp rafting phenomena and connectivity of nearshore communities around the coast of South Africa through development of a 3D spectral numerical coastal-flow model},
  pdfauthor={R Coppin; C Rautenbach; AJ Smit},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% https://github.com/rstudio/rmarkdown/issues/337
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

@@

\title{Quantifying passive kelp rafting phenomena and connectivity of nearshore
communities around the coast of South Africa through development of a 3D
spectral numerical coastal-flow model}
\author{R Coppin\footnote{University of the Western Cape} \and C Rautenbach\footnote{South African Weather Service} \and AJ Smit\footnote{University of the Western Cape}}
\date{}

@@

\begin{figure}
\centering
\includegraphics{thesis_chapter_1_files/figure-latex/surface elevation figures-1.pdf}
\caption{\label{fig:surface elevation figures}The definition of a wave
in a time record of the surface elevation with downward zero-crossings
(upper panel) and upward zero-crossings (lower panel).}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{thesis_chapter_1_files/figure-latex/Hs_Tp_surface_elevation figure-1.pdf}
\caption{The definition of wave height and wave period in a time record
of the surface elevation (the wave is defined with downwrd
zero-crossings.}
\end{figure}

@@

\hypertarget{linear-wave-theory}{%
\section{Linear wave theory}\label{linear-wave-theory}}

@@

The trajectory of passively drifting objects on the sea surface is
influenced by multiple factors, such as water currents, atmospheric
wind, wave motion, wave induced currents, gravitational force and
buoyancy force. To complicate matters, the previously mentioned factors
do not act independently but instead influence one another. Furthermore,
the gravitational and buoyancy forces on the object are determined by
the objects shape. Therefore, all these factors need to be taken into
account when modelling trajectory of drifting objects. Given the local
wind, surface current, and the shape and buoyancy of the object are
known, it is possible to estimate trajectory by the equation
\[V_{drift}=V_{curr} + V_{rel}\]

where \(V_{curr}\) represents the current velocity relative to the
earth, and \(V_{rel}\) represents the object drift velocity relative to
the ambient water (Hackett, Breivik, and Wettre 2006). Ocean currents
are determined by two components: the surface current (including the
effects of Ekman drift, baroclinic motion, tidal and inertial currents)
and Stokes drift induced by waves. The assumption made is \(V_{curr}\)
influences all floating objects in the same manner, regardless of shape
or size. Therefore, \(V_{curr}\) is equated with the surface current
obtained from a numerical ocean model which has been parameterised on
the wind velocity (Hackett, Breivik, and Wettre 2006). The effects of
\(V_{rel}\) on a floating object is driven by wind and wave forces which
is dependent on the shape and size of the floating object. In some cases
the effects of wind or wave forces may be neglibible and only one aspect
needs to be included or investigated seperately. Therefore, \(V_{rel}\)
can be seperated into wind and waves into two sub-components which
allows the equation to be adpated to suite the particluar situation
(Hackett, Breivik, and Wettre 2006). For example, previous research has
shown that the effects of waves on drifting objects are negligible when
the length of the object is less than the wave length; and increase
significantly when the lengths are approximately the same (Grue and
Biberg 1993; Hodgins and Hodgins 1998). Furthermore, current velocity
\(V_{curr}\) affects on drifting objects is considered to be negligible
in this approach and ideally should be taken into account. The
literature identifies waves as the primary influencer of trajectory for
larger objects while wind is regarded as the primary influencer for
smaller objects. The effects of wind on object drift trajectory is
complex and is ofcourse dependent on the exposed portion of the object
to the wind. The effect of wind on trajectory is often referred to as
leeway.

\hypertarget{leeway}{%
\subsection{Leeway}\label{leeway}}

An objects motion relative to the wind is sometimes referred to as
``windage factors'' or ``leeway drift'', and is difficult to accurately
and empirically describe. The difficulty is due to accurate measurements
needed for the current velocity, wind velocity, wave height and wave
direction. The appropriate approaches to measuring the afprmentioned
parameters have been detailed by ({\textbf{???}}) and is dependent on
the objects size and shape as well as the quality and quantity of data
available to the modeler. Additionally, these factors do not act
independently and instead influence one another, which compounds the
complexity of calculating wind effects on drifting objects. To overcome
this challenge, measurements of current and wind velocity, and wave
height and direction are measured along the objects trajectory for a
wide range of conditions.

@@

Important to note is that the effects of wind and waves on the
trajectory of a drifting object may be influenced by the object's
orientation. This has been noted in modelling drift trajectories of
ships (Allen and Plourde 1999; Hackett, Breivik, and Wettre 2006), and
should be incorporated into the leeway drift component. The direct
(\(V_l\)) and indirect (\(V_s\) and \(V_w\)) effects do not act
independently and instead influence one another. with regards to this,
Griffin, Oke, and Jones (2017) recommend incorporating an effective
windage factor which would represent the combined effects of Stokes
Drift, leeway drift and wave forces. Although a common approach, this
may not be necessary in all circumstances particularly if individual
effects of factors are being investigated. As mentioned previously, the
magnitude of effects of wind and waves on an objects trajectory is
depedent on the surface area/volume exposed to either influencer, which
in turn is dependent on the bouyancy of the object. The object's
buoyancy may also change over time due to other influencers such as
erosion and epibiont biomass.

@@

The buoyancy is often a fixed value within a model and does not alter
over time, such as @?? and @?? as examples. This may be because the
effects of epibiont biomass on ships and wreckage is not regarded as a
significant influencer. However, for smaller objects this may not be the
case. This may be particulary true for living drifters such as
macroalgae which are able to grow themselves, as well as provide a
habitat for epiphytes and bryozoans which increase in biomass overtime.
The buoyancy of objects at sea has been shown to be influeced by
epibiont load (Hobday 2000b; {\textbf{???}}; Graiff et al. 2016; Macaya
et al. 2016), which reduces drift times by reducing buoyancy with
increasing biomass, which ultimatley leads to the object sinking. The
bouyancy of the drifting object is determined by the growth rate of
epiphytic species while drifting, which in turn will be depndent on
environmental factors such as temperature and light which will vary
along the objects trajectory Therefore, buoyancy should be parameterised
by temporal epibiont biomass load.

Archimedes principles dictates that the buoyant force exerted on a small
object is equal to the weight of the displaced fluid and acts in the
opposite direction to the acceleraton vector. The net wieght is the
weight of the object less the weight of a equal volume of fluid. A
similar result can be achieved by using density of a object such that
the ``net density'' is the density of the object less the density of the
of the fluid. This can be expressed as:

\[weight=pgV_s\] \(Density=p_s-p_f\)

This approach can also be applied to acclerating objects submerged in a
fluid, in which case, the angular velocity must be considered.
Acceleration is usually considered in short-term forecasts for large
objects such as shipping containers or airplane debris. This does not
fall within the scope of this thesis and therefore will not be
considered.

@@

\begin{figure}
\centering
\includegraphics{thesis_chapter_1_files/figure-latex/BF_1-1.pdf}
\caption{\label{fig:BF_1}Conceptual model of bouyancy characteristics.}
\end{figure}

@@

\hypertarget{drag-parameters}{%
\subsection{Drag parameters}\label{drag-parameters}}

Drag is the force an object experiences when moving in a fluid due to
the rate in change of momentum (Vogel 2020). The drag around the surface
of an object due to the viscosity of the fluid is known as `surface
drag'. The drag induced by a quantity of fluid moving around the object
is known as pressure drag/form drag/inertial drag. In general, large
objects moving quickly tend to induce larger drag forces, however, this
is not always the case. In some instances, the flow around the object
changes from laminar to turbulent which ultimately reduces overall drag.
(Maybe insert small section on why turbulent flow reduces overall drag
on an object).

The type of flow, laminar or turbulent, is determined by fluid density
(p), flow speed (U), object size (l) and fluid viscosity (v). These
factors can be used to calculate the Reynolds number which is used to
determine the type of flow. The expression used can be dependent on
factors used such as whether the fluid is Newtonian or non-Newtonian and
object shape.

@@

Dispersal is recognised as important driver of biodiversity, composition
and structure of ecological systems in the marine environment (Bernardes
Batista et al. 2018; Helmuth, Veit, and Holberton 1994; Highsmith 1985;
MacArthur and Wilson 2001; Jackson and Sax 2010). These processes are
largely dependent on currents for dispersal of migrant populations which
ultimately promote connectivity of marine ecosystems (Bernardes Batista
et al. 2018; Mackas, Denman, and Abbott 1985, 1985; Zakas et al. 2009).
This is particularly true for organisms which lack pelagic larvae, and
are reliant on other modes of transport to reach new habitats through
passive modes of dispersal from anthropogenic and natural sources.
Anthropogenic sources include marine litter and structural debris while
examples of organic sources include pumice and macroalgae in the form of
kelp-rafts. Kelp-rafts have been identified as an important mode of
passive dispersal of marine organisms, such as invertebrates {[}Helmuth,
Veit, and Holberton (1994); whichmann2012{]}, epiphytes (Macaya et al.
2016), grazers (Nikula et al. 2010), invasive species (Lewis, Riddle,
and Smith 2005) and macroalgae themselves (Macaya et al. 2005) which
either make landfall or are transported offshore where they eventually
sink. A kelp-raft consists of an entanglement of macroalgae (one or
multiple species) which have been dislodged from the benthic environment
through hydrodynamic forces (mostly storms) and are positively buoyant
by means of gas-filled pneumatocysts and/or reproductive organs. Kelps
are capable of travelling vast distances (Fraser, Nikula, and Waters
2011) and are considered important dispersal vectors in temperate
latitudes, such as the Southern California Bight (Hobday 2000a, 2000b).

Kelp-raft abundance has been shown to vary temporally and spatially in
the ocean and around coastlines, which is dependent on seasonal growth
patterns which ultimatley influences overall biomass of seaweed. For
example, studies investigating the dispersal patterns of
\emph{Sargassum} in the West Pacific show increased abundance during
growth seasons (spring and summer) as more biomass is available to
fragment (Deysher and Norton 1981; Kingsford 1992). Temporal variability
may also be related to seasonal storm frequency where more storms in a
particular season relate to an increase in kelp-raft abundance. For
example a study by Kingsford (1995) investigated the contribution of
\emph{Macrocystis pyrifera} rafts to habitat complexity in pelagic
environments. The authors found that the abundance of \emph{M. pyrifera}
rafts increased during seasons where storms occurred more frequently.

The trajectory of kelp rafts are determined by a combination of
oceanographic conditions and biological processes. Since kelp float,
evidence points largely to wind driven surface currents as the main
driver, although the relative importance of wind versus surface currents
is largely unknown. For example a study by Harrold and Lisin (1989) used
radio-trackers on both natural and artificial \emph{Macrocystis
pyrifera} kelp rafts to investigate seasonal trajectory patterns. The
results showed that tracked kelp rafts were mostly deposited nearby the
source population and that seasonal wind direction was the primary
driver of trajectory for kelp rafts submerged roughly \textasciitilde0.5
- 1m below the surface Harrold and Lisin (1989). However, the authors
did note the influence of surface currents for some of their experiments
when winds relaxed; which suggests that surface currents may play a
larger role in determining trajectory for kelp rafts less exposed to
wind or when prevailing winds are low. Other authors have identified the
West Wind Drift as important driver of trajectory of floating \emph{M.
pyrifera} {[}Fraser, Nikula, and Waters (2011); {]}

@@

@@

\leavevmode\hypertarget{ref-Fraser2011}{}%
Fraser, C I, R Nikula, and J M Waters. 2011. ``Oceanic rafting by a
coastal community.'' \emph{Proceedings of the Royal Society B:
Biological Sciences} 278 (1706): 649--55.

@@

@@

\leavevmode\hypertarget{ref-helmuth1994}{}%
Helmuth, B, R R Veit, and R Holberton. 1994. ``Long-distance dispersal
of a subantarctic brooding bivalve (Gaimardia trapesina) by
kelp-rafting.'' \emph{Marine Biology} 120 (3): 421--26.
\url{https://doi.org/10.1007/BF00680216}.

@@

@@

\leavevmode\hypertarget{ref-Vogel2020}{}%
Vogel, Steven. 2020. ``Life in Moving Fluids: The Physical Biology of
Flow-Revised and Expanded Second Edition.''
